---
title: "p8105_hw1_xy2672.Rmd"
author: "Xirui Yang"
output: github_document
date: "2025-09-21"
---

```{r, echo = FALSE, message = FALSE}
library(moderndive)
library(tidyverse)
```

## Problem 1：

Packages already installed in console, now loading data:

```{r}
data("early_january_weather")
```

### Dataset description:
The variables included in the dataset are:  
``r names(early_january_weather)``.  
The early january weather dataset has **`r nrow(early_january_weather)`** rows and **`r ncol(early_january_weather)`** columns.  
The mean temperature of the early january weather is **`r round(mean(pull(early_january_weather, temp)), 2)`**.

Now creating the scatterplot of temperature vs. time_hour:

```{r}
weather_scatterplot = ggplot(early_january_weather, aes(x = time_hour, y = temp, color = humid)) + geom_point()
```

The scatterplot repeatedly shows a daily cycle. Based on the scatterplot, it is clear that the temperature generally increases from Jan 1 to Jan 14. In addition, temperature tends to be higher at the noon, and lower at night. Observations from nighttime and early morning are generally more humid than those around noon as shown in the scatterplot.

Saving the scatterplot:

```{r, message = FALSE}
ggsave("Weather_scatterplot.png", plot = weather_scatterplot)
```

## Problem 2:

Set seed for reproducibility
```{r}
set.seed(1234)
```

Creating the dataframe:

```{r}
p2_df = tibble(
  vec_num = rnorm(10),
  vec_logical_gt0 = vec_num > 0,
  vec_char = c("a","b","c","d","e","f","g","x","y","z"),
  vec_factor = factor(c("red","red","red","green","green","green","green","blue","blue","blue"))
)
```

